---
title: "Problem5"
date: "2022-11-14"
output: pdf_document
---

## Problem 5: Computational Finance - Modelling Stock prices

Following piece of code download the prices of TCS since 2007

```{r}
library(quantmod)
getSymbols('TCS.NS')
tail(TCS.NS)
```
Plot the adjusted close prices of TCS
```{r}
plot(TCS.NS$TCS.NS.Adjusted)
```

**Download the data of market index Nifty50**. The Nifty 50 index indicates how the over all market has done over the similar period.
```{r}
getSymbols('^NSEI')
tail(NSEI)
```
Plot the adjusted close value of Nifty50
```{r}
plot(NSEI$NSEI.Adjusted)
```


### Log-Return 
We calculate the daily log-return, where log-return is defined as
$$
r_t=\log(P_t)-\log(P_{t-1})=\Delta \log(P_t),
$$
where $P_t$ is the closing price of the stock on $t^{th}$ day.

```{r}
TCS_rt = diff(log(TCS.NS$TCS.NS.Adjusted))
Nifty_rt = diff(log(NSEI$NSEI.Adjusted))
retrn = cbind.xts(TCS_rt,Nifty_rt) 
retrn = na.omit(data.frame(retrn))
plot(retrn$NSEI.Adjusted,retrn$TCS.NS.Adjusted
     ,pch=20
     ,xlab='Market Return'
     ,ylab='TCS Return'
     ,xlim=c(-0.18,0.18)
     ,ylim=c(-0.18,0.18))
grid(col='grey',lty=1)
```

+ Consider the following model:

$$
r_{t}^{TCS}=\alpha + \beta r_{t}^{Nifty} + \varepsilon,
$$
where $\mathbb{E}(\varepsilon)=0$ and $\mathbb{V}ar(\varepsilon)=\sigma^2$.

1. Estimate the parameters of the models $\theta=(\alpha,\beta,\sigma)$ using the method of moments type plug-in estimator discussed in the class.
```{r}
ExpY = mean(retrn[,1])
ExpX = mean(retrn[,2])
VarY = var(retrn[,1])
VarX = var(retrn[,2])
CovXY = cov(retrn[,1],retrn[,2])
beta_hat_moments = CovXY/VarX
alpha_hat_moments = ExpY - (beta_hat_moments*ExpX)
sigma_hat_moments = sqrt((sum((retrn[,1]-alpha_hat_moments-beta_hat_moments*retrn[,2])^2))/length(retrn[,1]))
```
The optimal values of parameters are:
```{r}
c(alpha_hat_moments,beta_hat_moments,sigma_hat_moments)
```

2. Estimate the parameters using the `lm` built-in function of `R`. Note that `lm` using the OLS method.

```{r}

fit = summary(lm(retrn[,1]~retrn[,2], data=retrn))
beta_hat_OLS = fit$coef[2,1]
alpha_hat_OLS = fit$coef[1,1]
sigma_hat_OLS = fit$sigma
```
The optimal values of parameters are:
```{r}
c(alpha_hat_OLS,beta_hat_OLS,sigma_hat_OLS)

```
3. Fill-up the following table

Parameters | Method of Moments | OLS
-----------|-------------------|-----
$\alpha$   |    0.000461       |  0.000461  
$\beta$    |    0.743697       |  0.743697
$\sigma$   |    0.016184       |  0.016188
\newpage

4. If the current value of Nifty is 18000 and it goes up to 18200. The current value of TCS is Rs. 3200/-. How much you can expect TCS price to go up?
```{r}
TCS_return = alpha_hat_moments + beta_hat_moments * log(18200/18000)
TCS_new = 3200*exp(TCS_return)
```
New TCS price would be expected to be Rs.
```{r}
TCS_new
```
