---
title: "Problem2"
date: "2022-11-14"
output: pdf_document
---
## Problem 2 : Simulation Study to Understand Sampling Distribution

**Part A**
Suppose $X_1,X_2,\cdots,X_n\stackrel{iid}{\sim} Gamma(\alpha,\sigma)$, with pdf as
$$
f(x | \alpha,\sigma)=\frac{1}{\sigma^{\alpha}\Gamma(\alpha)}e^{- x/\sigma}x^{\alpha-1},~~~~0<x<\infty,
$$
The mean and variance are $E(X)=\alpha\sigma$ and $Var(X)=\alpha\sigma^2$. Note that `shape = ` $\alpha$ and `scale = ` $\sigma$.
1. Write a `function` in `R` which will compute the MLE of $\theta=\log(\alpha)$ using `optim` function in `R`. You can name it `MyMLE`
```{r}
My_MLE <- function(x){
      ll_gamma1 <- function(p){
        ll_gamma(p,x)
      }
      z <- optim(par=c(1,1),ll_gamma1) 
      log(z$par[1])
}
```

2. Choose `n=20`, and `alpha=1.5` and `sigma=2.2`
     (i) Simulate $\{X_1,X_2,\cdots,X_n\}$ from `rgamma(n=20,shape=1.5,scale=2.2)`
     (ii) Apply the `MyMLE` to estimate $\theta$ and append the value in a vector
     (iii) Repeat the step (i) and (ii) 1000 times
     (iv) Draw histogram of the estimated MLEs of $\theta$.
     (v) Draw a vertical line using `abline` function at the true value of $\theta$.
     (vi) Use `quantile` function on estimated $\theta$'s to find the 2.5 and 97.5-percentile points.
     
```{r}
library(stats)
set.seed(109)
vector = c()
gamma_simulation_MLE <- function(N,sim_size,alpha,sigma){
  for (j in 1:sim_size)
  {
    x<-rgamma(N,shape=alpha,scale=sigma)
    
    ll_gamma <- function(p,x){
      alpha<-p[1]
      sigma<-p[2]
      N*alpha*log(sigma)+N*log(gamma(alpha))-(alpha-1)*sum(log(x))+sum(x)/sigma
    }
    
    My_MLE <- function(x){
      ll_gamma1 <- function(p){
        ll_gamma(p,x)
      }
      z <- optim(par=c(1,1),ll_gamma1) 
      log(z$par[1])
    }
    vector <- c(vector, My_MLE(x))
  }
  hist(vector)
  abline(v=log(alpha),col='red')
  cat("The 2.5 and 97.5 percentile points are:")
  quantile(vector,probs=c(0.025,0.975))
}
gamma_simulation_MLE(20,1000,1.5,2.2)


```
3.  Choose `n=40`, and `alpha=1.5` and repeat the (2).
```{r}
gamma_simulation_MLE(40,1000,1.5,2.2)
```
4.  Choose `n=100`, and `alpha=1.5` and repeat the (2).
```{r}
gamma_simulation_MLE(100,1000,1.5,2.2)
```
5. Check if the gap between 2.5 and 97.5-percentile points are shrinking as sample size `n` is increasing?

Yes the gap between the 2.5 and 97.5 percentile points gets smaller as n increases.
 
